---
title: "HW11: Data Visualization I(cont"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load Libraries
```{r}
library(tidyverse)
library(here)
library(GGally)
library(patchwork)
```

### Import Data, Remove Columns, Remove Row, and Create Dummy Variable

```{r}
crabs_raw <- read.csv(here("Data", "aegla.crabs.csv"))
```

### View Imported Dataset
```{r}
crabs_raw
```

### Remove columns, remove row, create dummy variable
```{r}
crabs_data <- crabs_raw %>%
  select(Width, Depth, Flow, pH, TDS, Cond, N, Ni, Nt, Phos, Mg, Ca) %>%
  drop_na() %>%
  mutate(box_group="Crabs", row_order = 1:nrow(.))
```

### View Results
```{r}
view(crabs_data)
```

### Exploring width outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Width)) +
  geom_boxplot(width = 0.25)

```

## 1 outlier found in width data

### Exploring depth outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Depth)) +
  geom_boxplot(width = 0.25)
```

## 2 outliers found in depth data

### Explore flow outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Flow)) +
  geom_boxplot(width = 0.25)
```

## Multiple possible outliers

### Explore pH outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = pH)) +
  geom_boxplot(width = 0.25)
```

## 2 possible outliers found in pH data

### Explore TDS outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = TDS)) +
  geom_boxplot(width = 0.25)+
labs(x = NULL,
       y = "Total Dissolved Solids")
```

## No outliers

### Explore Cond outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Cond)) +
  geom_boxplot(width = 0.25)+
  labs(x = NULL,
       y = "Conductivity")
```
## No outliers

### Explore N outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = N)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
       y = "Nitrogen")
```

## Two outliers

### Explore Ni outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Ni)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
       y = "Nitrate")
```

## No outliers

### Explore Nt outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Nt)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
       y = "Nitrite Concentrations")
```
## Multiple possible outliers

### Explore Phos outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Phos)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
       y = "Phosphate")
```
## Two outliers

### Explore Mg outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Mg)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
       y = "Magnesium")
```
## No outliers

### Explore Ca outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Ca)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
       y = "Calcium concentrations")
```
## No outliers

## Overall interpretation is that the variables which most likely contain outliers are width, depth, pH, phosphate, and nitrogen. The variables flow and nitrate contained multiple outlier points







































               
               
               
               
               
               
               
               
               