---
title: "HW11: Data Visualization I(cont"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load Libraries
```{r}
library(tidyverse)
library(here)
library(GGally)
library(patchwork)
```

### Import Data, Remove Columns, Remove Row, and Create Dummy Variable

```{r}
crabs_raw <- read.csv(here("Data", "aegla.crabs.csv"))
```

### View Imported Dataset
```{r}
crabs_raw
```

### Remove columns, remove row, create dummy variable
```{r}
crabs_data <- crabs_raw %>%
  select(Width, Depth, Flow, pH, TDS, Cond, N, Ni, Nt, Phos, Mg, Ca) %>%
  drop_na() %>%
  mutate(box_group="Crabs", row_order = 1:nrow(.))
```

### View Results
```{r}
view(crabs_data)
```

### Exploring width outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Width)) +
  geom_boxplot(width = 0.25)

```

## 1 outlier found in width data

### Exploring depth outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Depth)) +
  geom_boxplot(width = 0.25)

```

## 2 outliers found in depth data

### Explore flow outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Flow)) +
  geom_boxplot(width = 0.25)

```

## Multiple possible outliers

### Explore pH outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = pH)) +
  geom_boxplot(width = 0.25)

```

## 2 possible outliers found in pH data

### Explore TDS outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = TDS)) +
  geom_boxplot(width = 0.25)+
labs(x = NULL,
     y = "Total Dissolved Solids")

```

## No outliers

### Explore Cond outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Cond)) +
  geom_boxplot(width = 0.25)+
  labs(x = NULL,
       y = "Conductivity")

```
## No outliers

### Explore N outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = N)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
        y = "Nitrogen")

```

## Two outliers

### Explore Ni outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Ni)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
        y = "Nitrate")

```

## No outliers

### Explore Nt outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Nt)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
        y = "Nitrite Concentrations")

```
## Multiple possible outliers

### Explore Phos outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Phos)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
        y = "Phosphate")

```
## Two outliers

### Explore Mg outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Mg)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
        y = "Magnesium")

```
## No outliers

### Explore Ca outliers
```{r}
crabs_data %>% 
  ggplot(aes(x = box_group, 
             y = Ca)) +
  geom_boxplot(width = 0.25)+
   labs(x = NULL,
        y = "Calcium concentrations")

```
## No outliers

## Overall interpretation is that the variables which most likely contain outliers are width, depth, pH, phosphate, and nitrogen. The variables flow and nitrate contained multiple possible outlier points. The variables TDS, Cond, Ni, Mg, and Ca contained no possible outliers. The data must be further explored to determine if possible outliers are truly outliers.

### Cleveland Plot #1

```{r}
clev_width <- crabs_data %>% 
  ggplot(aes(x = Width,
             y = row_order)) +
  geom_point() +
  labs(x = "Width",
       y = "Order of the data")

```

```{r}
clev_width
```

## Most likely what was reported as a possible outlier was just an upper value limit.


### Cleveland Plot #2
```{r}
clev_depth <- crabs_data %>% 
  ggplot(aes(x = Depth,
             y = row_order)) +
  geom_point() +
  labs(x = "Depth",
       y = "Order of the data")

```

```{r}
clev_depth
```

## Depth of 5 is most likely an outlier

### Cleveland Plot #3

```{r}
clev_pH <- crabs_data %>% 
  ggplot(aes(x = pH,
             y = row_order)) +
  geom_point() +
  labs(x = "pH",
       y = "Order of the data")
clev_pH

```

## Most likely what was reported as a possible outlier was just an upper value limit.



### Cleveland Plot #4

```{r}
clev_Nitrogen <- crabs_data %>% 
  ggplot(aes(x = N,
             y = row_order)) +
  geom_point() +
  labs(x = "Nitrogen",
       y = "Order of the data")
clev_Nitrogen

```

## Most likely what was reported as a possible outlier was just an upper value limit.

### Patchwork of Cleveland Plots

```{r}
clev_width+clev_depth+clev_pH+clev_Nitrogen
```

### Constructing Histograms
## Histogram #1: Width
```{r}
h1 <- crabs_data %>%
  ggplot() +
  geom_histogram(aes(x = Width),
                 binwidth = 1.0,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(1, 14, by = 2))
h1

```

## Histogram #2: Depth
```{r}
h2 <- crabs_data %>%
  ggplot() +
  geom_histogram(aes(x = Depth),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(0, 5, by = 0.5))
h2

```

## Histogram #3: pH
```{r}
h3 <- crabs_data %>%
  ggplot() +
  geom_histogram(aes(x = pH),
                 binwidth = 0.1,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(6, 8, by = 0.25))
h3

```

### Constructing Density Plots

## Width Density Plot #1
```{r}
dens1 <- crabs_data %>%
  ggplot() +
  geom_density(aes(x = Width))+
  scale_x_continuous(breaks = seq(1,  14, by = 2))
dens1

```


## Depth Density Plot #2
```{r}
dens2 <- crabs_data %>%
  ggplot() +
  geom_density(aes(x = Depth))+
  scale_x_continuous(breaks = seq(0,  5, by = 0.5))
dens2

```

## pH Density Plot #3
```{r}
dens3 <- crabs_data %>%
  ggplot() +
  geom_density(aes(x = pH))+
  scale_x_continuous(breaks = seq(6,  8, by = 0.25))
dens3

```

### Patchwork for Histograms and Density Plots #1
```{r}
h1+dens1
```

### Patchwork for Histograms and Density Plots #2
```{r}
h2+dens2
```

### Patchwork for Histograms and Density Plots #3
```{r}
h3+dens3
```


### Using ggpairs to investigate a relationship among all variables

```{r}
crabs_data %>% ggpairs(columns = 1:12)
```

### We made it, but lost as to its interpretation. 




               
               
               
               
               
               
               
               
               