---
title: "HW05 Part 1: Graphing Biological Data"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
install.packages("tidyverse")
library(tidyverse)
```

#### Plot 1

```{r}
data(trees)
```

```{r}
?trees
```

```{r}
ggplot(data = trees)+
  geom_point(mapping = aes(x=Height, y=Girth))
```

Looking at the graph provided, as height increases, so does girth. 

#### Plot 2

```{r}
data("lynx")
```

```{r}
?lynx
```

```{r}
year <- c(1821:1934)
```

```{r}
class(lynx)
```

```{r}
pelts <- as.vector(lynx)
```


```{r}
lynx_pelts <- data.frame(year, pelts)
```

```{r}
ggplot(data = lynx_pelts)+
  geom_smooth(mapping = aes(x= year, y= pelts, color= "maroon"))
```

There is a decline in the number of pelts recorded between 1840-1880. 

#### Plot 3

```{r}
data("iris")
```

```{r}
?iris
```

```{r}
ggplot(data = iris, mapping = aes(x= Petal.Length, y= Petal.Width))+ labs(y= "Petal Width", x= "Petal Length") + geom_point(mapping = aes(color= Species, shape= Species, point= 2))
```

Species setosa yields the smallest petals, while virginica yields the argest.

#### Plot 4

```{r}
data("msleep")
```

```{r}
?msleep
```

```{r}
ggplot(data = msleep, mapping = aes(x= vore, y= sleep_rem))+
  geom_violin(fill = "gray30")
```
Herbivores and omnivores spend the largest time in rem stage sleep.

#### Plot 5

```{r}
ggplot(data = msleep, mapping = aes(x= vore, y= sleep_rem))+
  geom_violin(fill = "gray70")
```

Herbivores and omnivores spend the largest time in rem stage sleep.
Violin plot is a mirrored density plot.  Fill allows the densities to be displayed on a grayscale, ranging from a solid fill to a mere outline.

#### Plot 7 because there is not a plot 6. 

```{r}
ggplot(data = msleep, mapping = aes(x= order, y= sleep_total))+
  geom_boxplot()+
  coord_flip()
```
Coordinate flip was helpful with this dataset because the order labels are lengthy. 

#### Plot 8

```{r}
ggplot(data = msleep, mapping = aes(x= conservation, y= awake))+ 
  labs(y= "Awake", x= "Conservation")+ 
  geom_boxplot()+ geom_jitter()+
  geom_point(mapping = aes(color= conservation))+guides(col=guide_legend("Conservation"))+
  coord_flip()
```

The use of coord_flip allows the data to be more spread out instead of compacted. Therefore, making it easier to read. 

#### Plot 9

```{r}
ggplot(data = msleep, mapping = aes(x= sleep_total, y= bodywt))+
  geom_point(mapping = aes(color=vore))
```

Herbivores have larger body weights compared to the other vore types.

```{r}
ggplot(data = msleep, mapping = aes(x= sleep_total, y= brainwt))+
  geom_point(mapping = aes(color=conservation))+facet_wrap(~vore, ncol = 3)
```

Insectivores tend to have longer total sleep cycles.