---
title: "HW 08: More Data Wrangling"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 5: Data transformation (PART 1)

### Section 5.5: Add new variables with `mutate()`

## Import and view data from flights

```{r}
flights_sml <- select(flights, year:day, ends_with("delay"), distance, air_time)
```

## Add new columns to the end of the dataset

```{r}
mutate(flights_sml,gain = dep_delay - arr_delay,speed = distance / air_time * 60)
```

## Refer to the columns created

```{r}
mutate(flights_sml,gain = dep_delay - arr_delay,hours = air_time / 60,gain_per_hour = gain / hours)
```

## Only keepng the new variables by using the function 'transmute ()'

```{r}
transmute(flights,gain = dep_delay - arr_delay, hours = air_time / 60, gain_per_hour = gain / hours)
```

### Section 5.5.1: Useful creation functions

## Use modular arithmetic

```{r}
transmute(flights,dep_time,hour = dep_time %/% 100,minute = dep_time %% 100)
```
## Use offsets lead() and lag()

```{r}
(x <- 1:10)
```

```{r}
lag(x)
```

```{r}
lead(x)
```

## Use cumulative and rolling aggregates 

```{r}
x
```

```{r}
cumsum(x)
```

```{r}
cummean(x)
```

## Ranking functions 
```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
```

```{r}
min_rank(desc(y))
```

## Ranking functions with variants

```{r}
row_number(y)
```

```{r}
dense_rank(y)
```

```{r}
percent_rank(y)
```

```{r}
cume_dist(y)
```


### Section 5.6: Grouped summaries with `summarise()`

## Collapse data frame to a single row

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

## Change the unit of analysis from teh complete dataset to individual groups

```{r}
by_day <- group_by(flights, year, month, day) 
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

### 5.6.1 Combining multiple operations with the pipe

## Explore the relationship between the distance and average delay for each location

```{r}
by_dest <- group_by(flights, dest) 
delay <- summarise(by_dest, count = n(), dist = mean(distance, na.rm = TRUE), delay = mean(arr_delay, na.rm = TRUE))
delay <- filter(delay, count > 20, dest != "HNL")
```


## Create a scatterplot from the data

```{r}
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

## Group the same data using pipe

```{r}
 delays <- flights %>% 
  group_by(dest) %>% 
  summarise(count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)) %>% 
  filter(count > 20, dest != "HNL")
```

### 5.6.2: Missing values

## Not setting an argument will result in this... 

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```
## Using the 'na.rm' function to remove the missing values

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```
## Remove cancelled flights

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

## View the flights
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```
### Skipping 5.6.3 (I've had enough baseball to last me for another year after this hot season)

### 5.6.4: Useful summary functions

## Using 'mean(x)' and 'median(x)' function

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0]) # the average positive delay
  )
```
## Measures of rank using quantiles

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time))
```
## Measures of position

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time))
```
## Filter variables

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

## Counts

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```

## Counts with simple helper (Use only if you want to do is count)

```{r}
not_cancelled %>% 
  count(dest)
```

## Count using a weight variable 

```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```

## Counts and proportions of logical values

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_prop = mean(arr_delay > 60))
```

### 5.6.5: Grouping by multiple variables

## Group by multiple vairables to easily roll up a dataset

```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
```

```{r}
(per_month <- summarise(per_day, flights = sum(flights)))
```

```{r}
(per_year  <- summarise(per_month, flights = sum(flights)))
```
### 5.6.6: Ungrouping

## Remove grouping using the function 'ungroup()'

```{r}
daily %>% 
  ungroup() %>%          
  summarise(flights = n()) 
```

### Section 5.7: Grouped mutates (and filters)

## Worst members of each group

```{r}
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```

## All groups bigger than a threshold

```{r}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```

## Standardise to compute per group metrics.

```{r}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```



















