---
title: "HW07 Part 1:Data Import and Wrangling 1"
author: "Alissa Seal"
date: "`r format(sys.time(), '%d %B %Y)`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

```{r}
file_path <- "../data" 
```

### Limpets 

Load Limpets Data

```{r}
limpets_data <- read_csv(file.path(file_path, "limpets.csv"))
```
The limpets data is tidy. 

Use the column types argument

```{r}
col_types = "ncc"
```

Make the sites column an ordered factor

```{r}
limpets_data$site <- factor(limpets_data$Site, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"))
```

Check sorting

```{r}
limpets_data
```

Make a boxplot
```{r}
ggplot(data = limpets_data, mapping = aes(x= site, y= Length)) +
  geom_boxplot(aes(color=Protected))+ labs(x="Site", y= "Length (mm)")
```

DP and VBG sites have outliers. The protected sites are home to larger limpets.

### Roseate Terns

The data is not tidy.

```{r}
tern_data <- read_tsv(file.path(file_path, "roseate_terns.txt"))
```


Remove sites with missing count data

```{r}
tern_data %>% filter(complete.cases(.))
```
Remove sites with missing data

```{r}
complete_terns <- tern_data %>% filter (complete.cases(.))
```

Tidy Data

```{r}
tern_2 <- complete_terns %>%
  pivot_longer(AR:JC, names_to = "Sites", values_to = "Populations")
```

```{r}
tern_2
```

Tern data graph

```{r}
tern_2 %>%
  ggplot() +
  geom_line(aes(x= year, y = Populations, group= Sites, color= Sites))
```

Several sites increased in population, including, AR, AL, AC, BL, BC, JL, JC. The decreasing sites would include BR and JR. Some lines have breaks in them becuase of the missing data 'NA' that was removed. 

### Backlip Abalone

# Chunk 1: 

Import abalone data, remove the first coloumn, and make a boxplot of height differences among the three types. 

```{r}
abalone_data <- read_csv(file.path(file_path, "abalone.csv"))
select(abalone_data, -(X1))
ggplot(data = abalone_data) +
  geom_boxplot(mapping = aes(x= Type, y= Height))
```

# Chunk 2: 

Make a scatterplot to determine if height correlates with rings. 

```{r}
ggplot(data = abalone_data)+
  geom_point(mapping = aes(x= Rings, y= Height))
```

The outliers are not old trees. They are young trees. 

# Chunk 3

Filter data to remove outliers. Create an ordered factor. Graph a scatterplot.

```{r}
abalone_data_2 <- filter(abalone_data, Height <= 0.35)
factor(abalone_data_2$Type, levels = c("I", "F", "M"))
ggplot(data = abalone_data_2) +
  geom_point(mapping = aes(x = Rings, y = Height, color = Type))
```

The new scatterplot shows the mature individualls generally have greater height than immature individuals. However, not all mature male and female individuals are larger than immature.

# Chunk 4

Create a plot with zero height individuals removed.

```{r}
abalone_data_2[abalone_data_2$Height == 0, 'Height'] <- NA
ggplot(data = abalone_data_2) +
  geom_point(mapping = aes(x = Height, y = Rings, color = Type))
```
# Chunk 5

```{r}
ggplot(data = abalone_data_2) +
  geom_point(mapping = aes(x = Rings, y = Diameter, color = Type))
```


```{r}
ggplot(data = abalone_data_2) +
  geom_point(mapping = aes(x = Rings, y = LongestShell, color = Type))
```

```

### Darters

```{r}
darters_data <- read_fwf(here("data", "darter.txt"), fwf_widths(c(3, 11, 1, 1, 1, 3, 4, 4, 4, 3, 3, 2, 2)), comment = '#')
named_darter <- darters_data %>% 
  rename(ID = X1, Species = X2, Riffle = X3, Age = X4, Sex = X5, Length = X6, Depth = X7, Velocity = X8, Temperature = X9, Major_Sub = X10, Minor_Sub = X11, Major_Type = X12, Minor_Type = X13)
```









