---
title: "HW09: Wrangling II"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

```{r}
library(here)
```

```{r}
library(knitr)
```

```{r}
library(smatr)
```

### Source my_functions
```{r}
source(here("scripts","my_functions.R"))
```

### Import darters.csv and remove species tetrazonum and zonale
```{r}
darters_data <- read_csv(here("Data", "darters.csv"))
darters_data_filtered <- darters_data %>%
  filter(species=="blennioides"|species=="caeruleum"|species== "spectabile")
```

### Sum majsub and minsub to form new column. Change riffle values. Change length units. Change sex values.
```{r}
wrangled_darters_data <- mutate(darters_data_filtered, total_substrate=majsub+minsub) %>%
  mutate(riffle=ifelse(riffle==1,
                       "Riffle 1",
                       "Riffle 2")) %>%
  mutate(length=length*10) %>%
  mutate(sex=ifelse(sex=="f",
                    "Female",
                    "Male"))
```

### Summarize the mean length, depth, and velcoity. Group by species and riffle. 
```{r}

spec_riffle_sep_wrang_darters_data <- group_by(wrangled_darters_data, species, riffle)
summary_wrangled_data <- summarize(spec_riffle_sep_wrang_darters_data, length=mean(length, na.rm = TRUE), depth=mean(depth, na.rm = TRUE), velocity=mean(velocity, na.rm = TRUE))
```

### Use kable to make a table.

```{r}
kable(summary_wrangled_data)
```

### Plot 1: Summarized darters data

```{r}
ggplot(data = wrangled_darters_data) + 
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```
### Add geom_point() to add mean length

```{r}
ggplot(data = summary_wrangled_data) +
  geom_point(mapping = aes(x=species, y=length), shape=15)
stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

### Rotate axis, separate in to 2 riffle, add labs()

```{r}
ggplot(data = summary_wrangled_data) +
  geom_point(mapping = aes(x=species, y=length), shape=15)+
  facet_wrap(~riffle) +
  coord_flip() +
  labs(y = "length", x = NULL)
stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

# Riffle 2 contains the longest average blennioides, but among the other species, the individuals in riffle one are longer

### Darters Plot 2

## Added gray scale and first stat_summary()

```{r}
ggplot(data = summary_wrangled_data) +
  geom_point(mapping = aes(x=species, y=length), color= "gray65")+
  facet_wrap(~riffle) +
  coord_flip() +
  labs(y = "length", x = NULL)
stat_summary(
    fun.y = median,
    geom = "point",
    size = 2.2
  )
```



### Darters Plot 2

## Add second stat_summary() function to the graph

```{r}
ggplot(data = summary_wrangled_data, aes(x= species, y = length)) +
  geom_point(color= "gray65")+
stat_summary(
    fun.y = median,
    geom = "point",
    size = 2.2
  ) +
stat_summary(fun.y = mean, geom = "point", size = 3,color = "maroon")+
  facet_wrap(~riffle) +
  coord_flip() +
  labs(y = "length", x = NULL)
```

```{r}
library(tidyverse)
library(smatr)
```

### Stress and Coritcosterones In Birds

### Import feathers.csv and change decimal mark

```{r}
feathers_data <- read_csv(here("Data", "feathers.csv"), locale = locale(decimal_mark = ","))
```
## Renaming Feather Columns

```{r}
feathers_data_2 <- feathers_data %>%
  rename("capture_date" = "Capture Date", "day" = "Date",  "site" = "Site", "urban_score" = "Urba Score", "cortF" = "CortF ng/mg", "cortB" = "CortF ng/mm", "sex" = "Sex", "smi" =  "SMI", "tarsus" = "Tarsus", "mass" = "Mass", "basal_cort" = "Bas_Cort", "stress_cort" = "Stress_Cort")
```

## View feathers_data_2

```{r}
feathers_data_2
```

## Calculate mean tarsus length

```{r}
mean_tarsus <- mean(feathers_data_2$tarsus, na.rm = TRUE)
```

## Use mutate() to form new columns
```{r}
mutated_feathers_data_2 <- mutate(feathers_data_2, log_mass=log(mass), log_tarsus= log(tarsus))
```

## Use sma() function to calculate the standardized major axis regression


```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = mutated_feathers_data_2)
```

## Apply summary function
```{r}
summary(major_axis)
```

## Apply coef() function

```{r}
coef(major_axis)
```

## Store slope value in b

```{r}
slope <- coef(major_axis)
b <- slope['slope']
```

## Use mutate to add new SMI column

```{r}
feathers_data_2_w_SMI <- mutate(feathers_data_2,new_smi = scaled_mass(mass,tarsus,b))
```



















