---
title: "HW10: Data Visualization I"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Install and load libraries

```{r}
library(tidyverse)
```

```{r}
library(here)
```

```{r}
library(GGally)
```

```{r}
library(patchwork)
```

### Import Sparrows Data

```{r}
sparrows_raw <- read_tsv(here("data", "sparrows.txt"))

```

## Check
```{r}
sparrows_raw
```

### Wrangle Sparrow Data

```{r}
sparrows <- sparrows_raw %>% 
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))

```

## Check
```{r}
view(sparrows)
```

### Figure 2: Outliers

## Run Code 
```{r}
plot_box<- sparrows %>% 
  ggplot(aes(x = box_group, y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL, y = "Wing length (mm)")

```
## Run Cleveland dot plot

```{r}
plot_cleveland <- sparrows %>%
  ggplot(aes(x= wingcrd,y=row_order))+
  geom_point()+
  labs(x="Wing length (mm)", y="Order of the date")

```

## Display plots
```{r}
plot_box+plot_cleveland
```
## Add coord_flip()

```{r}
plot_box+plot_cleveland+coord_flip()
```

### Figure 3: Outliers

## culmen plot
```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


```

## nalospi plot
```{r}
p_nalospi <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


```

## wt plot
```{r}
p_wt <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


```

## wingcrd plot
```{r}
p_wingcrd <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

```

## tarsus plot
```{r}
p_tarsus <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


```

## head plot
```{r}
p_head <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

```

## Six Variable Plot
```{r}
p_culmen+p_nalospi+p_wt+p_wingcrd+p_tarsus+p_head
```

## The variables nalospi, tarsus, and head appear to have outlier data

### Figure 5: Distibution

```{r}
h1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
```


## Faceted histogram

```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1)
```

## Patchwork

```{r}
h1+h2
```

## Recreate figure using density

## Create plot 1

```{r}
dens1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_density(aes(x = wt))+
  scale_x_continuous(breaks = seq(14,  28, by = 2))

```

## Create plot 2
```{r}
dens2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_density(aes(x = wt, fill=Month), alpha=0.50) +
  scale_x_continuous(breaks = seq(14, 28, by = 2))

```

## Patching density plots together

```{r}
dens1+dens2
```
### Figure 10:Relationships

