---
title: "HW14: Maps"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Import Data
```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

### Part 1: Outline Maps

## Outline of the 48 contiguous states
```{r}
usa <- map_data(map= "state")
```

## Outline of just Missouri
```{r}
us_missouri <- map_data (map = "state",
                         region = "missouri")

```

## Outline of Alabama, Florida, Mississippi, and North Carolina
```{r}
us_other <- subset(x= usa, 
                   subset= region %in% c("alabama",
                                         "florida",
                                         "mississippi",
                                         "north carolina"))

```

```{r}
str(usa)
```

## Map of 4 non-Missouri states
```{r}
ggplot()+
  geom_polygon(data = us_other, 
               aes(x=long, 
                   y= lat,
                   group=group),
               fill= "gray")

```

## Add Missouri with emphasis 
```{r}
ggplot() +
geom_polygon(data = us_other,
             aes(x = long,
                 y = lat,
                 group = group),
             fill = "gray") +
geom_polygon(data = us_missouri,
             aes(x = long, y = lat,
                 group = group),
             fill = "black")

```

## Outline of all 48 states
```{r}
ggplot() +
geom_polygon(data = us_other,
             aes(x = long,
                 y = lat,
                 group = group),
             fill = "gray") +
geom_polygon(data = us_missouri,
             aes(x = long, y = lat,
                 group = group),
             fill = "black") +
geom_polygon(data = usa,
             aes(x=long,
                 y = lat,
                 group = group),
             fill = NA,
             color = "black")

```
## Add the filled states first so that the black outline of all
## States are overlaid on top
```{r}
us_map <- ggplot() +
geom_polygon(data = us_other,
             aes(x = long,
                 y = lat,
                 group = group),
             fill = "gray") +
geom_polygon(data = us_missouri,
             aes(x = long, y = lat,
                 group = group),
             fill = "black") +
geom_polygon(data = usa,
             aes(x=long,
                 y = lat,
                 group = group),
             fill = NA,
             color = "black") +
  theme_map() +
coord_fixed(1.3)
us_map

```

## Missouri Map
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))
mo_map <- ggplot() +
geom_polygon(data = missouri, aes(x=long, y = lat, group = group), fill = NA, color = "black") +
geom_polygon(data = mo_scott, aes(x = long, y = lat, group = group), fill = "black") +
theme_map() +
coord_fixed(1.3)
mo_map

```

## us_map and a mo_map side by side
```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5,1))

```


### Outline Maps: My Turn

## Outline of the 48 contiguous states
```{r}
usa <- map_data(map= "state")

```

## Outline of just Arkansas
```{r}
us_arkansas <- map_data (map = "state",
                         region = "arkansas")

```


## U.S. map should highlight in black the state of your birth
```{r}
us_map <- ggplot() +
geom_polygon(data = us_arkansas,
             aes(x = long, y = lat,
                 group = group),
             fill = "black") +
geom_polygon(data = usa,
             aes(x=long,
                 y = lat,
                 group = group),
             fill = NA,
             color = "black") +
theme_map() +
coord_fixed(1.3)
us_map

```

## State map highlighting the county of your birth
```{r}
arkansas <- map_data("county", "arkansas")
ar_greene <- subset(arkansas, subregion %in% c("greene"))
ar_map <- ggplot() +
geom_polygon(data = arkansas, aes(x=long, y = lat, group = group), fill = NA, color = "black") +
geom_polygon(data = ar_greene, aes(x = long, y = lat, group = group), fill = "black") +
theme_map() +
coord_fixed(1.3)
ar_map

```

## Outline other counties
```{r}
ar_other_counties <- subset(x= us_arkansas, subregion %in% c("arkansas","sharp"))

```



## Two additional counties in gray
```{r}
arkansas <- map_data("county", "arkansas")
ar_greene <- subset(arkansas, subregion %in% c("greene"))
ar_other_counties <- subset(arkansas, subregion %in% c("arkansas", "sharp"))
ar_map_counties <- ggplot() +
geom_polygon(data = ar_other_counties, aes(x=long, y = lat, group = group), fill = "gray") +
geom_polygon(data = arkansas, aes(x=long, y = lat, group = group), fill = NA, color = "black") +
geom_polygon(data = ar_greene, aes(x = long, y = lat, group = group), fill = "black") +
theme_map() +
coord_fixed(1.3)
ar_map_counties

```

### Part 2: Choropleth Maps

## Load Libraries
```{r}
library(sf)
library(tidyverse)
```

## Import skin cancer data
```{r}
skin_cancer_raw <-read_delim(file = here::here("data", "skin_cancer.txt"),
delim = " ",
skip = 7,
trim_ws = TRUE)

```
## Wrangle the Data

```{r}
skin_cancer <- skin_cancer_raw %>%
filter(State != "Wash,D.C.") %>%
mutate(State = case_when(
State == "NorthCarolina" ~ "North Carolina",
State == "SouthCarolina" ~ "South Carolina",
State == "NorthDakota" ~ "North Dakota",
State == "SouthDakota" ~ "South Dakota",
State == "NewHampshire" ~ "New Hampshire",
State == "NewJersey" ~ "New Jersey",
State == "NewMexico" ~ "New Mexico",
State == "MewYork" ~ "New York", # Data has MewYork typo
State == "RhodeIsland" ~ "Rhode Island",
State == "WestVirginia" ~ "West Virginia",
TRUE ~ State))
skin_cancer

```

## Calculate mortality rate to the mean rate
```{r}
mean_mort <- mean(skin_cancer$Mort)
```
```{r}
skin_cancer_adj <- skin_cancer %>%
mutate(relative_mort = Mort - mean_mort)
mean_mort

```

# Trim Alaska, Hawaii, and territories from the shapefile.
```{r}
lower_48 <- c("Alabama","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida", "Georgia", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Mryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming")

```

## Use filter() to keep only states in the states object that listed in the lower_48 vector

```{r}
states <- st_read(here::here("data/shape_files", "cb_2017_us_state_500k.shp"),
stringsAsFactors = TRUE)
states

```

## Create a tibble witht eh state shapes to plot
```{r}
states_df <- states %>%
  dplyr::filter(NAME %in% lower_48)
states_df

```
## FIPS States
```{r}
str(states_df$STATEFP)
```

## Capture the FIPS levesl from states_df
##Use to factor the FIPS and states names in the next step

```{r}
state_levels <- levels(states_df$STATEFP)
name_levels <- levels(states_df$NAME)
```

## Import he state FIPS, remove an unneeded column, rename the FIPS column to STATEFP, then use mutate() to change the STATEFP column to an unordered factor

```{r}
library(tidyverse)
fips <- read_csv(here::here("data", "state_fips.csv")) %>% 
  dplyr::select(-abbr)%>%
  rename(STATEFP = FIPS) %>% 
  mutate(STATEFP = factor(STATEFP, levels = state_levels))
fips

```

## Use left_join to merge the data in a single file

```{r}
skin_cancer_adj <- left_join(skin_cancer_adj, fips)
```

```{r}
states_df <- left_join(states_df, skin_cancer_adj)
```

## Plot 
```{r}
ggplot(states_df) +
  geom_sf(aes(fill = relative_mort)) +
  scale_fill_viridis_c(name = "Skin Cancer,\nRelative Mortality",
                       option = "inferno") +
  coord_sf(crs = st_crs(5070)) +
  theme_bw() +
  theme(legend.position = "bottom")

```

### Chropleth Map: My Turn
## Import Data
```{r}
lyme_disease_raw <- read_csv(here::here("data", "lyme_disease.csv"),skip = 1, na = c("-"))
lyme_disease_raw

```

### Add New 2017 Column, Remove Columns, Use pivot_longer(), Group by State, and Summarize Mean
```{r}
lyme_disease_tidy <- lyme_disease_raw %>%
  mutate(`2017` = `2017_Confirmed` + `2017_Probable`)%>%
  dplyr::select(State:`2016`, `2017`)%>%
  pivot_longer("2007":"2017",names_to = "Year",values_to = "Cases") %>%
  group_by(State) %>% 
  summarise(Mean_Cases = mean(Cases),Log_Mean = log10(Mean_Cases))
lyme_disease_tidy

```

## Merging the fips data frame to Lyme Disease Data

```{r}
lyme_disease_tidy <-left_join(lyme_disease_tidy, fips)
states_df <-left_join(states_df, lyme_disease_tidy)

```
## Plot Lyme Data
```{r}
ggplot(states_df) +
  geom_sf(aes(fill = Log_Mean)) +
  scale_fill_viridis_c(name = "Lyme Disease \n Mean Cases",
                       option = "plasma", labels = c("<1", "10", "100", "1000", "5000")) +
  coord_sf(crs =st_crs(5070)) +
  theme_bw() +
  theme(legend.position = "bottom")

```

### Part 3: Dot Distribution Map

### Load Libraries
```{r}
library(raster)
library(tidyverse)
library(ggthemes)
library(ggmap)

```

### Global Vars
```{r}
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(9618973)

```


### Bigfoot Data
```{r}
bigfoot <- read_csv(here::here("data", "bigfoot.csv"))

bigfoot <- bigfoot %>% 
  filter(long >= min_long & long <= max_long,
         lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus") %>% 
  dplyr::select(name, long, lat)

```
### Bear Data
```{r}
bears <- read_csv(here::here("data", "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)

```

```{r}
tail(both_species)
```

### Terrain Map
```{r}
base = get_stamenmap(bbox = c(min_long, 
                              min_lat, 
                              max_long, 
                              max_lat),
                     zoom = 4, 
                     maptype = "terrain-background")



ggmap(base)

```
### Terrain Map with Bigfoot location
```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))

```
### Terrain Map with Bears Location
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))

```

### Map with wooded areas bears prefer
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) +
  scale_shape_manual(values = c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")

```

### My Part of Part 3


### Global Vars
```{r}
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(1513373)

```


### Bigfoot Data
```{r}
bigfoot <- read_csv(here::here("data", "bigfoot.csv"))

bigfoot <- bigfoot %>% 
  filter(long >= min_long & long <= max_long,
         lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus") %>% 
  dplyr::select(name, long, lat)

```
### Bear Data
```{r}
bears <- read_csv(here::here("data", "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)

```

```{r}
tail(both_species)
```

### Terrain Map
```{r}
base = get_stamenmap(bbox = c(min_long, 
                              min_lat, 
                              max_long, 
                              max_lat),
                     zoom = 4, 
                     maptype = "terrain-background")



ggmap(base)

```
### Terrain Map with Bigfoot location
```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))

```
### Terrain Map with Bears Location
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))

```

### Map with wooded areas bears prefer
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) +
  scale_shape_manual(values = c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")

```

### Part 4: Bathymetry

### Load Libraries
```{r}
library(tidyverse)
library(marmap)
```

### Global Vars
```{r}
min_long <- -170
max_long <- 164
min_lat  <- -42
max_lat  <- -8
```

## Conversion
```{r}
earthquakes <- quakes %>% 
   mutate(depth = depth * 3280.84)
```


## Auto read saved file
```{r}
fiji <- getNOAA.bathy(lon1 = min_long,
                      lon2 = max_long,
                      lat1 = min_lat,
                      lat2 = max_lat,
                      antimeridian = TRUE,
                      keep = TRUE)
```
## Autoplot without the .bathy ext

```{r}
autoplot.bathy(fiji, 
         geom = c("raster", "contour"),
         size = 0.1,
         na.rm = TRUE)
```

