---
---
title: "HW13: Strings and Things"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load Libraries 
```{r}
library(tidyverse)
library(here)
library(ggthemes)
```

### Part1: DNA Strings

### Import beta_gliadin data

```{r}
beta_gliadin_raw <- read_lines(here("Data", "beta_gliadin.txt"), skip = 1)
beta_gliadin_raw

```

### Replace whitespace and digits with nothing

```{r}
beta_gliadin_cleaned <- str_replace_all(beta_gliadin_raw, "\\d", "") %>%
  str_replace_all("\\s", "")
beta_gliadin_cleaned

```

### Collapse the strings

```{r}
beta_gliadan_collapsed <- str_c(beta_gliadin_cleaned, collapse = "")
beta_gliadan_collapsed

```

### Convrt to uppercase
```{r}
beta_gliadan_upper <- str_to_upper(beta_gliadan_collapsed)
beta_gliadan_upper

```

### Return the length of string
```{r}
str_length(beta_gliadan_upper)

```

### Count letters that are not A, C, G, T
```{r}
str_count(beta_gliadan_upper, ".A.C.G.T")

```

### Count the possible start codons
```{r}
str_count(beta_gliadan_upper, "ATG")

```

### Count the possible stop codons
```{r}
str_count(beta_gliadan_upper, "TGA|TAA")

```

### Create caat_box and tata_box variables
```{r}
caat_box <- "CCAAT"
tata_box <- "CTATAA"

```

### Count how many times caat_box occurs in the DNA sequence
```{r}
str_count(beta_gliadan_upper, caat_box)

```

### Count how many times tata_box occurs in the DNA sequence
```{r}
str_count(beta_gliadan_upper, tata_box)

```

### Make a regex search string called caat_seq
```{r}
caat_seq <- "CCAAT.{100,150}ATG"

```

### Count how many times this sequence occurs
```{r}
str_count(beta_gliadan_upper, caat_seq)

```

### Create tata_seq search string
```{r}
tata_seq <- "CTATAA.{50,100}ATG"

```

### Count how many times this sequence occurs
```{r}
str_count(beta_gliadan_upper, tata_seq)

```

### Extract all caat_seq, determine the length of each sequence 
```{r}
caat_seq_all <- str_extract_all(beta_gliadan_upper, caat_seq, simplify = TRUE)
str_length(caat_seq_all)

```

### Use str_which() 
```{r}
str_which(caat_seq_all, tata_seq)

```

### str_which() would only show the position of the match, not the match itself

### Use str_subset() to print the entire match
```{r}
promotor <- str_subset(caat_seq_all, tata_seq)
promotor

```

### Determine the length of the match
```{r}
str_length(promotor)
```

### PART 2: ANTS

### Import Ants Data
```{r}
ants_raw <- read_csv(here("Data", "ants.csv"))
ants_raw

```


### Wrangle the data and select columns
```{r}
ants_data_wrangled <- ants_raw %>% 
  pivot_longer(names_to="species",values_to="abundance",matches("[A-Z].+\\..+", ignore.case = FALSE)) %>%
  select("year", "species", "abundance")
ants_data_wrangled

```

### Rename Species
```{r}
ants_mutated <- ants_data_wrangled %>% 
  mutate(species=str_replace(species, "(^[A-Z][a-z]{0,3}).*\\.([a-z]{1,4}).*", "\\1\\2"))
ants_mutated

```

### Create 2010 and 2011 levels, group by year and species and use sum() to add up the abundance of each species

```{r}
final_ants_data <- ants_mutated %>%
  mutate(year = factor(year, ordered = TRUE, levels = c(2010, 2011))) %>%
   group_by(year, species) %>%
   summarise(total = sum(abundance))
final_ants_data

```

### Plot the data
```{r}
ggplot(final_ants_data,aes(x = reorder(species,total),y = total),group = species)+
  geom_line()+
  geom_point(aes(shape = year),size = 2)+
  coord_flip()+
  labs(x = NULL, y = "Total Abundance", shape = "Year")+
  theme_minimal() 

```

### Part 3: Featuring Phlox Flowers

### Import and wrangle the first data set

```{r}
flower_size_raw <- read_csv(here("Data", "flower_size.csv")) 
flower_size_raw

```

### Filter out rows were flower number is not NA
```{r}
flower_data_filtered <- drop_na(flower_size_raw, "Flower number")
flower_data_filtered

```

### Select Columns
```{r}
wrangled_flower_data <- select(flower_data_filtered, "Species", "Corolla length (cm)", "Corolla width throat (cm)") %>%
  rename("species" = "Species", "cor_length" = "Corolla length (cm)", "throat_width" = "Corolla width throat (cm)")
wrangled_flower_data

```

### Use separate to separate genus and species
```{r}
separated_flowers <- wrangled_flower_data %>%
  separate(species, into = c("genus", "species"))
separated_flowers

```

### Group data by genus and species and use summarize to calculate the mean and max lengths an widths
```{r}
summarized_flower_data <- separated_flowers %>%
  group_by(genus, species) %>%
  summarize("mean_cor_length" = mean(cor_length, na.rm = TRUE), "mean_throat_width" = mean(throat_width, na.rm = TRUE), "max_cor_length" = max(cor_length, na.rm = TRUE), "max_throat_width"= max(throat_width, na.rm = TRUE))
summarized_flower_data
```


### Import Data
```{r}
pollinators_raw <- read_csv(here("Data", "flower_pollinators.csv"))
pollinators_raw

```

### Remove columns, remove rows, separate into genus and species
```{r}
pollinators <- pollinators_raw %>%
   select(-c(starts_with("Source"), "Color")) %>%
   filter(!is.na(Pollinator)) %>%
   separate(Species, into = c("genus", "species"))
pollinators

```

### Separate pollinator column
```{r}
syndrome <- pollinators %>%
  separate(Pollinator, into = "syndrome", sep = "(,| )")
syndrome

```

### Use inner_join to join the data sets
```{r}
phlox_data <- inner_join(summarized_flower_data, syndrome)
phlox_data
```
  
### Phlox Plot 1 with improved legend
```{r}
pholox_plot_1 <- phlox_data %>%
  mutate(syndrome=str_replace(syndrome, "(^be.*e$)","\\1s"))%>%
  ggplot()+
  geom_point(mapping = aes(x=mean_cor_length, y=mean_throat_width, color= syndrome))
pholox_plot_1

```
### Check Syndrome Counts
```{r}
phlox_data %>%
  group_by(syndrome) %>%
  count(syndrome)

```

### Phlox Plot 2, Create other inscets in legend
```{r}
phlox_data_2 <- phlox_data %>%
  mutate(syndrome=str_replace(syndrome,"(^be.*e$)", "\\1s"),
         syndrome=str_replace(syndrome,".+.?es", "other insects"))

phlox_data_2 %>%
  ggplot()+
  geom_point(mapping = aes(x=mean_cor_length, y=mean_throat_width, color=syndrome))

```

### Mutate Data
```{r}
mutated_phlox <- phlox_data_2 %>%
  mutate(syndrome_group = case_when(syndrome=="autogamous"~"Autogamous", syndrome=="bat"~"Mammal", syndrome=="hummingbird"~"Bird", TRUE~"Insect"))
mutated_phlox

```

### Plot data, Phlox plot 3
```{r}
mutated_phlox %>%
  ggplot()+
  geom_point(mapping = aes(x=mean_cor_length, y=mean_throat_width, color= syndrome_group))

```

### Improved Plot
```{r}
mutated_phlox %>%
  ggplot()+
  geom_point(aes(x=max_cor_length, y= max_throat_width, color=syndrome_group))+
  scale_color_viridis_d(type= "qual", palette= "inferno")+
  labs(x="Mean Length (cm)", y="Mean Width (cm)", color="Syndrome")
```















