---
---
title: "HW13: Strings and Things"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load Libraries 
```{r}
library(tidyverse)
library(here)
library(ggthemes)
```

### Part1: DNA Strings

### Import beta_gliadin data

```{r}
beta_gliadin_raw <- read_lines(here("Data", "beta_gliadin.txt"), skip = 1)
beta_gliadin_raw

```

### Replace whitespace and digits with nothing

```{r}
beta_gliadin_cleaned <- str_replace_all(beta_gliadin_raw, "\\d", "") %>%
  str_replace_all("\\s", "")
beta_gliadin_cleaned

```

### Collapse the strings

```{r}
beta_gliadan_collapsed <- str_c(beta_gliadin_cleaned, collapse = "")
beta_gliadan_collapsed

```

### Convrt to uppercase
```{r}
beta_gliadan_upper <- str_to_upper(beta_gliadan_collapsed)
beta_gliadan_upper

```

### Return the length of string
```{r}
str_length(beta_gliadan_upper)

```

### Count letters that are not A, C, G, T
```{r}
str_count(beta_gliadan_upper, ".A.C.G.T")

```

### Count the possible start codons
```{r}
str_count(beta_gliadan_upper, "ATG")

```

### Count the possible stop codons
```{r}
str_count(beta_gliadan_upper, "TGA|TAA")

```

### Create caat_box and tata_box variables
```{r}
caat_box <- "CCAAT"
tata_box <- "CTATAA"

```

### Count how many times caat_box occurs in the DNA sequence
```{r}
str_count(beta_gliadan_upper, caat_box)

```

### Count how many times tata_box occurs in the DNA sequence
```{r}
str_count(beta_gliadan_upper, tata_box)

```

### Make a regex search string called caat_seq
```{r}
caat_seq <- "CCAAT.{100,150}ATG"

```

### Count how many times this sequence occurs
```{r}
str_count(beta_gliadan_upper, caat_seq)

```

### Create tata_seq search string
```{r}
tata_seq <- "CTATAA.{50,100}ATG"

```

### Count how many times this sequence occurs
```{r}
str_count(beta_gliadan_upper, tata_seq)

```

### Extract all caat_seq, determine the length of each sequence 
```{r}
caat_seq_all <- str_extract_all(beta_gliadan_upper, caat_seq, simplify = TRUE)
str_length(caat_seq_all)

```

### Use str_which() 
```{r}
str_which(caat_seq_all, tata_seq)

```

### str_which() would only show the position of the match, not the match itself

### Use str_subset() to print the entire match
```{r}
promotor <- str_subset(caat_seq_all, tata_seq)
promotor

```

### Determine the length of the match
```{r}
str_length(promotor)
```

### PART 2: ANTS

### Import Ants Data
```{r}
ants_raw <- read_csv(here("Data", "ants.csv"))
ants_raw

```


### Wrangle the data and select columns
```{r}
ants_data_wrangled <- ants_raw %>% 
  pivot_longer(names_to="species",values_to="abundance",matches("[A-Z].+\\..+", ignore.case = FALSE)) %>%
  select("year", "species", "abundance")
ants_data_wrangled

```

### Rename Species
```{r}
ants_mutated <- ants_data_wrangled %>% 
  mutate(species=str_replace(species, "(^[A-Z][a-z]{0,3}).*\\.([a-z]{1,4}).*", "\\1\\2"))
ants_mutated

```

### Create 2010 and 2011 levels, group by year and species and use sum() to add up the abundance of each species

```{r}
final_ants_data <- ants_mutated %>%
  mutate(year = factor(year, ordered = TRUE, levels = c(2010, 2011))) %>%
   group_by(year, species) %>%
   summarise(total = sum(abundance))
final_ants_data

```

### Plot the data
```{r}
ggplot(final_ants_data,aes(x = reorder(species,total),y = total),group = species)+
  geom_line()+
  geom_point(aes(shape = year),size = 2)+
  coord_flip()+
  labs(x = NULL, y = "Total Abundance", shape = "Year")+
  theme_minimal() 

```




















