---
---
title: "HW13: Strings and Things"
author: "Alissa Seal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load Libraries 
```{r}
library(tidyverse)
library(here)
library(ggthemes)
```

### Part1: DNA Strings

### Import beta_gliadin data

```{r}
beta_gliadin_raw <- read_lines(here("Data", "beta_gliadin.txt"), skip = 1)
beta_gliadin_raw

```

### Replace whitespace and digits with nothing

```{r}
beta_gliadin_cleaned <- str_replace_all(beta_gliadin_raw, "\\d", "") %>%
  str_replace_all("\\s", "")
beta_gliadin_cleaned
```

### Collapse the strings

```{r}
beta_gliadan_collapsed <- str_c(beta_gliadin_cleaned, collapse = "")
beta_gliadan_collapsed
```

### Convrt to uppercase
```{r}
beta_gliadan_upper <- str_to_upper(beta_gliadan_collapsed)
beta_gliadan_upper
```

### Return the length of string
```{r}
str_length(beta_gliadan_upper)
```

### Count letters that are not A, C, G, T
```{r}
str_count(beta_gliadan_upper, ".A.C.G.T")
```

### Count the possible start codons
```{r}
str_count(beta_gliadan_upper, "ATG")
```

### Count the possible stop codons
```{r}
str_count(beta_gliadan_upper, "TGA|TAA")
```

